Emulsions Stabilized by Magnetic Ellipsoidal Particles: A Lattice Boltzmann Study

### 2025-04-11 chapters/lit_review_new.tex
Changes 1. Rearranged introduction, lit review and methods sections

This strong attachment energy makes particles highly effective at stabilizing emulsions by becoming irreversibly trapped at fluid-fluid interfaces. 
\cite{ngai_particle-stabilized_2015}Once adsorbed, these 
particles reduce interfacial tension and act as physical barriers that prevent the coalescence of droplets, thereby enhancing emulsion stability. Unlike molecular surfactants, 
which are dynamic and can desorb under certain conditions, colloidal particles remain anchored at the interface due to the significant energy required to detach them. 
\cite{ngai_particle-stabilized_2015}This mechanism 
is especially useful in systems such as Pickering emulsions, where the interfacial coverage and wettability of the particles determine the type and stability of the emulsion formed. 
\cite{ngai_particle-stabilized_2015,velankar_non-equilibrium_2015}
As such, understanding particle behavior at interfaces is essential for controlling the structure and longevity of particle-stabilized systems. \cite{ngai_particle-stabilized_2015}
For particles that are smooth and simple and at an interface of oil and water, the attachment energy at the fluid interface can be calculated from a force 
balance between the force imparted from the particle on each fluid and the fluids between one another, known as the Young-Dupre equation.

\begin{equation}
    \cos{\theta_c} = \frac{\sigma_{po} - \sigma_{pw}}{\sigma_{ow}}
\end{equation}

Where $\theta_c$ is the contact angle, $\sigma_{po}$ is the surface tension between particle and oil, $\sigma_{pw}$ is the surface tension between particle and water and
$\sigma_{ow}$ is the surface tension between oil and water. The particle is considered to be preferentially wet by water or hydrophilic if $\theta_c < 90 ^{\circ}$ or 
hydrophobic if $\theta_c > 90 ^{\circ}$. The particle is said to be neutrally wetting if $\theta_c = 90 ^{\circ}$.

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{../figures/literature_review/particle_at_interface.png }
    \caption{Particle schematic at an interface demonstrating the various quantities of interest. Reprinted Figure 2.1 with permission from
             Particle-Stabilized Emulsions and Colloids: Formation and Applications; Ngai, T., Bon, S. A. F., Eds.; RSC soft matter series; 
             Royal Society of Chemistry, RSC Publishing Cambridge, 2015}
    \label{fig:pieranski_model}
\end{figure}

The energy of the system before and after particle adsorption is different and can be characterized for a simple system with parameters shown in Figure \ref{fig:pieranski_model},

\begin{equation}
    \Delta E_{IW} = E_I - E_w
\end{equation}

Where $\Delta E_{I}$ is the energy of the system when the particle is in the interface and $\Delta E_W$ is the energy of the system when the particle is completely in the water phase.
We define $\Delta E_{I}$ as

\begin{equation}
    E_I = A_w \sigma_{pw} + A_o\sigma_{po} + A_{I_1}\sigma_{ow} + V_w P_w + V_o P_o
\end{equation}

Where $A_w$ and $A_o$ are the surface areas of the particle in the water and oil phases respectively, $V_w$ and $V_o$ are the volumes of water and oil phases respectively and 
$P_w$ and $P_o$ are the pressures of water and oil phases respectively. $A_{I_1}$ is the area of the interface with the particle. 
A similar equation can be written for the energy of the water phase, $E_w$

\begin{equation}
    E_w = (A_w + A_o)\sigma_{pw} + A_{I_2}\sigma_{ow} + (V_w + V_o)P_w
\end{equation}

$A_{I_2}$ is the area of the interface without the particle. The change in the interface energy upon adsorption of the particle can be computed to obtain

\begin{equation}
    \Delta E_{IW} = A_o(\sigma_{po} - \sigma_{pw}) - A_p\sigma_{ow} + V_o(P_o - P_w)
\end{equation}

Substituting the Young-Laplace law for a spherical droplet, $P_w - P_o = \frac{2\sigma_{ow}}{R_d}$ where $R_d$ is the droplet size, for $(P_o - P_w)$ and the Young-Dupre contact angle
expression, the following is obtained

\begin{equation}
    \Delta E_{IW} = \sigma_{ow}(A_o\cos{\theta_c} - A_p - \frac{2 V_o}{R_d})
\end{equation}

A similar construction can be done for the oil phase to obtain

\begin{equation}
    \Delta E_{IW} = -\sigma_{ow}(A_w\cos{\theta_c} + A_p - \frac{2 V_o}{R_d})
\end{equation}

After making appropriate geometrical substitutions for $V_o$ and $A_p$, the Piersanski model is recovered,

\begin{equation}
    \Delta E_{ads} = \pi r^2 \sigma_{ow} (1 - |\cos{\theta_c}|)^2
\end{equation}

From this expression, the adsorption energy can be seen to depend on the contact angle, surface tension and contact area of the particles. This indicates that larger, neutrally wetting
particles adsorbing onto interfaces with high surface tension have a high driving force for adsorption. \cite{ngai_particle-stabilized_2015,reeves_particle-size_2015} This adsorption process 
can be orders of magnitude larger than the thermal energy of the system, calculated as $k_b T$ where $T$ is the temperature of the system and $k_b$ is the Boltzmann constant. This results in 
irreversible adsorption of particles at interfaces. \cite{binks_pickering_2001}

In addition to the adsorption of particles on the interface, the contact angle of the particles determine the dispersed phase of an emulsion and its microstructure. 
\cite{velankar_non-equilibrium_2015}
Hydrophillic particles tend to prefer to form oil in water emulsions while hydrophobic particles prefer to form water in oil emulsions. \cite{ngai_particle-stabilized_2015}
This occurs as the particles preferentially wet one phase over another, causing the interface to bend as the emulsion forms. Neutrally wetting particles are preferred in the formation of 
bijels as no preferential curvature is imparted onto the bijels, fulfilling one of the requirements of a minimal surface in that the mean curvature must be zero. \cite{jinnai_interfacial_2001}

@@

The thermodynamic equilibrium of a binary fluid mixture can be described by its Gibbs free energy, given as \(\Delta G = \Delta H - T \Delta S\), where \(\Delta H\) is the enthalpy change, 
\(T\) is temperature, and \(\Delta S\) is the entropy change. The onset of phase separation occurs at the binodal line, defined where \(\frac{dG}{dx} = 0\), with \(x\) representing the composition. 
Although the system is metastable, it remains uniform until fluctuations give rise to small nuclei. The free energy of each nucleus reflects a 
balance between the bulk free energy gain from phase separation and the interfacial energy penalty, expressed as \(\Delta G = \frac{4}{3}\pi r^3 \Delta g_v + 4 \pi r^2 \sigma\), where \(r\) is 
the nucleus radius, \(\Delta g_v\) is the free energy change per unit volume, and \(\sigma\) is the interfacial tension. 
\cite{thanh_mechanisms_2014}
Nuclei that exceed the critical radius \(r_c = \frac{2\sigma}{|\Delta g_v|}\) will grow, driving the system toward phase separation via nucleation and growth. \cite{thanh_mechanisms_2014}
This transformation process is commonly modeled by the Avrami equation, \(Y = 1 - \exp(-K t^n)\), where \(Y\) is the 
transformed fraction, \(K\) is a rate constant incorporating nucleation and growth kinetics, and \(n\) is the Avrami exponent, which reflects the dimensionality and mechanism of the transformation.
\cite{avrami_kinetics_1939}

At the binodal, the system is still metastable, meaning that unless perturbations are introduced to form nuclei, the mixture remains homogeneous. \cite{thanh_mechanisms_2014}
However, a deeper quench into the two-phase region leads to the spinodal line, defined by the stability criterion \(\frac{d^2G}{dx^2} = 0\). \cite{cahn_free_1958}
Within this region, the system is unstable to even infinitesimal composition fluctuations, and phase 
separation proceeds spontaneously without the need for nucleation. This process, known as spinodal decomposition, is characterized by the continuous and simultaneous growth of interconnected 
domains of both phases. \cite{cahn_free_1958} Unlike nucleation and growth, which relies on localized events, spinodal decomposition 
results in a bicontinuous structure driven by the system's attempt to minimize 
free energy through composition modulation. \cite{cahn_free_1958} The dynamics of spinodal decomposition were characterized by Cahn and Hilliard. \cite{cahn_free_1958} 
We first begin by defining the free energy of a regular solution, known as the bulk free energy.

@@

@@

As the library of available particle shapes continues to grow, it becomes increasingly important to understand how these anisotropic particles behave at fluid interfaces.
\cite{wu_recent_2016, cavallaro_curvature-driven_2011}
Unlike spherical particles, anisotropic particles introduce directional interactions and complex rotational dynamics which can significantly 
influence interfacial assembly and emulsion stability. \cite{read_dimerization_2020, davies_dipolar_2015,morgan_understanding_2013}
Their shape can dictate not only their orientation at the interface but also how they interact, and jam, 
ultimately affecting the structure and mechanical properties of the resulting material. \cite{hijnen_bijels_2015}
This is relevant for systems like bijels, where the precise arrangement of particles at the interface affects how the bijel james in place, generating new timescales or 
interfacial phenomena not characterized in bijels made with spheres. \cite{gunther_timescales_2014,hijnen_bijels_2015}

Anisotropic particles exhibit shape-dependent interfacial behavior due to their ability to induce complex multipolar capillary interactions. When adsorbed at a fluid-fluid interface, these 
particles distort the surrounding interface in order to maintain a mean curvature of zero, in accordance with the Young-Laplace equation. This distortion creates curvature gradients around the 
particle that interact with nearby particles or features on the interface, giving rise to anisotropic capillary forces. \cite{loudet_capillary_2005, cheng_shape-anisotropic_2013}. 
Computational simulations further support this complexity, showing that ellipsoidal particles can adopt a variety of metastable orientations 
depending on initial conditions and perturbations encountered during adsorption or rearrangement at the interface \cite{gunther_lattice_2013, newton_influence_2014}.

@@

When looking at arrays of ellipsoidal particles at interfaces, curvature gradients control the dynamics of the system. Figure \ref{fig:curvature_gradient} illustrates the effect that
Cavallaro and co-workers identified. They showed how curvature gradients created at the edges of a 
square rod caused directed assembly of cylinders, with the driving force strong enough to overcome gravity. \cite{cavallaro_curvature-driven_2011} Other works have introduced how these features are 
particle shape dependent with greater control over the curvature gradient providing a means to control particle self assembly. \cite{read_dimerization_2020, sharifi-mood_curvature_2015} Botto et al. 
identified that ellipsoidal particles prefer to arrange side to side state while cylinders prefer to arrange tip to tip with the particle geometric smoothness playing a role in this preference.
\cite{botto_capillary_2012} These differences in particle arrangement could open up new possibilities for tailoring interfacial assemblies and the subsequent differences in microstructure obtained in
a bijel.

Ellipsoidal particles have been used as stabilizers in bijels, resulting in bijels that have smaller domains than those made with spherical particles at the same volume fraction. This was identified to arise 
from the greater surface area to volume ratio of ellipsoidal particles compared to spherical particles. \cite{gunther_timescales_2014} The evolution of the microstructure is also distinct to bijels 
stabilized with spheres, demonstrating two timescales at early and later timesteps. \cite{gunther_timescales_2014} The short timescale is during formation of the bijel, when particles are adsorbed onto the
interface and rotate to align with the interface. At longer timescales, capillary interactions between particles lead to reordering of particles at the interface, leading to a different interfacial 
configuration and microstructure. \cite{gunther_timescales_2014}

@@

Rod like particles have also been used as stabilizers in experiments, identifying that the length scale of the pores of the bijels follows $ L \propto \frac{1}{\phi_p} $ even if the proportionality constant 
is different. Figure \ref{fig:rod_bijel_flippers} demonstrates bijels stabilized by rod like particles showing that particles flip out of the interface due to steric forces from adjacent particles. This
flipping can increase the resistance to shear due to capillary bridging between adjacent rods.
While the proportionality constant depends on particle geometry, this trend has been observed across various particle shapes 
\cite{hijnen_bijels_2015, madivala_exploiting_2009, gunther_timescales_2014, daware_emulsions_2015, loudet_capillary_2005, cheng_shape-anisotropic_2013}. 
The dynamics of bijels stabilized by ellipsoidal particles are also distinct from bijels stabilized by spheres, caused by reorientation of particle stabilizers at the interface. \cite{gunther_timescales_2014}

@@

Understanding the rheology of emulsions, particularly bijels, is essential because it directly impacts both their fabrication and practical use. \cite{haase_situ_2016}
Previous investigations have characterized that bijels are non-Newtonian fluids with a yield stress and shear thinning behavior. \cite{macmillan_rheological_2019}
A non-Newtonian fluid is one whose viscosity is not constant but varies with the applied shear rate or stress. In Newtonian fluids like water 
or air, there is a linear relationship between shear stress and shear rate, meaning viscosity remains the same regardless of how fast the fluid is sheared. \cite{mezger_rheology_2020} Non-Newtonian fluids, on 
the other hand, deviate from this behavior and may exhibit shear thinning, shear thickening, yield behavior or viscoelasticity with no fluid constrained to one behavior. \cite{mezger_rheology_2020}
Shear thinning and thickening here define that the change in viscosity is sub-linear and super-linear respectively, while yield behavior implies that the viscosity is non-zero when no shear 
is applied. \cite{mezger_rheology_2020} Viscoelasticity appears when the shear response of the fluid is dependent upon the frequency of the applied shear, indicating that the fluid has both liquid and solid like 
characteristics. \cite{mezger_rheology_2020}
These properties often appear in soft matter and multiphase systems where microstructure and particle interactions influence flow behavior.

@@

Suspensions and emulsions are examples of non-Newtonian fluids due to the presence of dispersed particles or droplets within a continuous phase. 
\cite{brader_nonlinear_2010, besseling_three-dimensional_2007, xu_relation_2013}  
In suspensions, the interaction  between solid particles—such as hydrodynamic forces, electrostatic repulsion, and steric hindrance—creates microstructural rearrangements under flow, leading to 
non-linear viscosity 
responses. \cite{brader_nonlinear_2010, besseling_three-dimensional_2007} Figure \ref{fig:suspension_shear} demonstrates the shear properties of an example suspension, demonstrating how shear thinning
and shear thickening can occur in suspensions. At low shear, particles may aggregate or form a network that resists deformation, while higher shear rates can break 
down these structures, resulting in shear thinning. In dense systems, particles may experience contact-induced resistance or jamming, causing shear thickening. \cite{brader_nonlinear_2010}

Emulsions behave similarly with droplet deformation, coalescence, and interactions 
among droplets generating time-dependent and non-linear responses to shear. When particles are used to stabilize emulsions (e.g., in Pickering emulsions), their interfacial rigidity 
adds further resistance to flow and deformation, contributing to complex, often viscoelastic, rheological behavior.
Madivala et al. showed that Pickering emulsions stabilized with rod-like particles had enhanced stability under shear, with greater resistance observed as particle aspect ratio increased 
\cite{madivala_exploiting_2009}. They identified this improvement to be due to improved interfacial coverage of ellipsoidal particles at interfaces and the improved elasticity of the interface,
allowing for the droplet to elongate more before failure. These improvements also extend to disc like particles with other investigations into the resistance to shear of emulsions stabilized by
graphene nanosheets, showing that they were able to provide better resistance to shear than spherical particles. \cite{imperiali_simple_2014} This was due to the graphene nanosheets overlapping one
another during shear, creating greater resistance to the applied shear and a higher apparent viscosity. In this case too, the overlap of graphene nanosheets was shown to result in greater interfacial
elasticity inferred through surface tension measurements of droplets with graphene stabilizers. \cite{sun_assembly_2013}

@@

Several theoretical models have been developed to describe the behavior of non-Newtonian fluids under constant shear. The schematic representation of these models is shown in
Figure \ref{fig:shear_theory}. The power law Model is commonly used to represent shear thinning or 
thickening, with the equation \(\sigma = k \dot{\gamma}^n\), where \(\sigma\) is shear stress, \(\dot{\gamma}\) is shear rate, \(k\) is the consistency index, and \(n\) is the flow index.
\cite{mezger_rheology_2020}
For \(n<1\), the fluid is shear thinning and for \(n>1\), shear thickening. \cite{mezger_rheology_2020}
The Bingham plastic model and Herschel-Bulkley model account for yield stress, with the latter being more general by 
including power-law behavior after yielding, defined as $\sigma = \sigma_y + k \dot{\gamma}^n$. \cite{mezger_rheology_2020}
Shear models that include viscoelasticity have also been developed such as the  Maxwell and Kelvin-Voigt 
models. \cite{mezger_rheology_2020}
They are used when the fluid exhibits both solid- and liquid-like characteristics under stress, incorporating time-dependent strain responses and relaxation phenomena. These models help 
predict and quantify the behavior of complex fluids in real-world applications. \cite{mezger_rheology_2020}

>>>>

@@

The thermodynamic equilibrium of a binary fluid mixture can be described by its Gibbs free energy, given as $G = \Delta H - T \Delta S$, where \(\Delta H\) is the enthalpy change, 
\(T\) is temperature, and \(\Delta S\) is the entropy change. The onset of phase separation occurs at the binodal line, defined where \(\frac{dG}{dx} = 0\), with \(x\) representing the composition. 
Although the system is metastable, it remains uniform until fluctuations give rise to small nuclei. The free energy of each nucleus reflects a 
balance between the bulk free energy gain from phase separation and the interfacial energy penalty, expressed as \(\Delta G = \frac{4}{3}\pi r^3 \Delta g_v + 4 \pi r^2 \sigma\), where \(r\) is 
the nucleus radius, \(\Delta g_v\) is the free energy change per unit volume, and \(\sigma\) is the interfacial tension \cite{thanh_mechanisms_2014}. 
Nuclei that exceed the critical radius \(r_c = \frac{2\sigma}{|\Delta g_v|}\) will grow, driving the system toward phase separation via nucleation and growth \cite{thanh_mechanisms_2014}. 
This transformation process is commonly modeled by the Avrami equation, \(Y = 1 - \exp(-K t^n)\), where \(Y\) is the 
transformed fraction, \(K\) is a rate constant incorporating nucleation and growth kinetics, and \(n\) is the Avrami exponent, which reflects the dimensionality and mechanism of the transformation
\cite{avrami_kinetics_1939} .

At the binodal, the system is still metastable, meaning that unless perturbations are introduced to form nuclei, the mixture remains homogeneous \cite{thanh_mechanisms_2014}.
However, a deeper quench into the two-phase region leads to the spinodal line, defined by the stability criterion \(\frac{d^2G}{dx^2} = 0\) \cite{cahn_free_1958}. 
Within this region, the system is unstable to even infinitesimal composition fluctuations, and phase 
separation proceeds spontaneously without the need for nucleation. This process, known as spinodal decomposition, is characterized by the continuous and simultaneous growth of interconnected 
domains of both phases \cite{cahn_free_1958}. Unlike nucleation and growth, which relies on localized events, spinodal decomposition 
results in a bicontinuous structure driven by the system's attempt to minimize 
free energy through composition modulation \cite{cahn_free_1958}. The dynamics of spinodal decomposition were characterized by Cahn and Hilliard \cite{cahn_free_1958} . 
We first begin by defining the free energy of a regular solution, known as the bulk free energy.

@@

When micron sized particles are adsorbed by interfaces the attachment energy is large enough that they can become irreversibly trapped at fluid-fluid interfaces \cite{ngai_particle-stabilized_2015}. 
Once adsorbed, these 
particles reduce interfacial tension and act as physical barriers that prevent the coalescence of droplets, thereby making them highly effective emulsion stabilizers. Unlike molecular surfactants, 
which are dynamic and can desorb under certain conditions, colloidal particles remain adsorbed at the interface due to the significant energy required to detach them \cite{ngai_particle-stabilized_2015}. 
This mechanism 
is especially useful in systems such as Pickering emulsions, where the interfacial coverage and wettability of the particles determine the type and stability of the emulsion formed 
\cite{ngai_particle-stabilized_2015,velankar_non-equilibrium_2015}.
As such, understanding particle behavior at interfaces is essential for controlling the structure and longevity of particle-stabilized systems \cite{ngai_particle-stabilized_2015}. 
For particles that are smooth and simple and at an interface of oil and water, the attachment energy at the fluid interface can be calculated the Pieranski model, 
$E_{ads} = \sigma A_{rm} (1 - \cos{\theta_c})^2$, where $E_{ads}$ is the free energy reduction upon particle adsorption, $\sigma$ is the surface tension between the two 
fluids, and $\theta_c$ is the contact angle of the stabilizing particle.

This expression is derived from an energy balance between the interactions between the interface and the 
fluid phases, with a schematic shown in Figure \ref{fig:pieranski_model},

\begin{figure}
    \centering
    \includegraphics[scale = 0.5]{../figures/literature_review/particle_at_interface.png }
    \caption{Particle schematic at an interface demonstrating the various quantities of interest. Reprinted Figure 2.1 from
             Particle-Stabilized Emulsions and Colloids: Formation and Applications; Ngai, T., Bon, S. A. F., Eds.; RSC soft matter series; 
             with permission Royal Society of Chemistry, RSC Publishing Cambridge, 2015 under license number 1597404-1}
    \label{fig:pieranski_model}
\end{figure}

From the Pieranski model, the adsorption energy can be seen to depend on the contact angle, surface tension and contact area of the particles. This indicates that larger, neutrally wetting
particles adsorbing onto interfaces with high surface tension have a high driving force for adsorption. \cite{ngai_particle-stabilized_2015,reeves_particle-size_2015} This adsorption process 
can be orders of magnitude larger than the thermal energy of the system, calculated as $k_b T$ where $T$ is the temperature of the system and $k_b$ is the Boltzmann constant. This results in 
irreversible adsorption of particles at interfaces \cite{binks_pickering_2001}.

In addition to the adsorption of particles on the interface, the contact angle of the particles determine the dispersed phase of an emulsion and its microstructure through the
Bancroft rule \cite{velankar_non-equilibrium_2015,ngai_particle-stabilized_2015}. 
Hydrophillic particles tend to prefer to form oil in water emulsions while hydrophobic particles prefer to form water in oil emulsions \cite{ngai_particle-stabilized_2015}.
This occurs as the particles preferentially wet one phase over another, causing the interface to bend as the emulsion forms. Neutrally wetting particles are preferred in the formation of 
bijels as no preferential curvature is imparted onto the bijels, fulfilling one of the requirements of a minimal surface in that the mean curvature must be zero \cite{jinnai_interfacial_2001}.

% from a force 
% balance between the force imparted from the particle on each fluid and the fluids between one another, known as the Young-Dupre equation.

% \begin{equation}
%     \cos{\theta_c} = \frac{\sigma_{po} - \sigma_{pw}}{\sigma_{ow}}
% \end{equation}

% Where $\theta_c$ is the contact angle, $\sigma_{po}$ is the surface tension between particle and oil, $\sigma_{pw}$ is the surface tension between particle and water and
% $\sigma_{ow}$ is the surface tension between oil and water. The particle is considered to be preferentially wet by water or hydrophilic if $\theta_c < 90 ^{\circ}$ or 
% hydrophobic if $\theta_c > 90 ^{\circ}$. The particle is said to be neutrally wetting if $\theta_c = 90 ^{\circ}$.

% \begin{figure}
%     \centering
%     \includegraphics[scale = 0.5]{../figures/literature_review/particle_at_interface.png }
%     \caption{Particle schematic at an interface demonstrating the various quantities of interest. Reprinted Figure 2.1 from
%              Particle-Stabilized Emulsions and Colloids: Formation and Applications; Ngai, T., Bon, S. A. F., Eds.; RSC soft matter series; 
%              with permission Royal Society of Chemistry, RSC Publishing Cambridge, 2015 under license number 1597404-1}
%     \label{fig:pieranski_model}
% \end{figure}

% The energy of the system before and after particle adsorption is different and can be characterized for a simple system with parameters shown in Figure \ref{fig:pieranski_model},

% \begin{equation}
%     \Delta E_{IW} = E_I - E_w
% \end{equation}

% Where $\Delta E_{I}$ is the energy of the system when the particle is in the interface and $\Delta E_W$ is the energy of the system when the particle is completely in the water phase.
% We define $\Delta E_{I}$ as

% \begin{equation}
%     E_I = A_w \sigma_{pw} + A_o\sigma_{po} + A_{I_1}\sigma_{ow} + V_w P_w + V_o P_o
% \end{equation}

% Where $A_w$ and $A_o$ are the surface areas of the particle in the water and oil phases respectively, $V_w$ and $V_o$ are the volumes of water and oil phases respectively and 
% $P_w$ and $P_o$ are the pressures of water and oil phases respectively. $A_{I_1}$ is the area of the interface with the particle. 
% A similar equation can be written for the energy of the water phase, $E_w$

% \begin{equation}
%     E_w = (A_w + A_o)\sigma_{pw} + A_{I_2}\sigma_{ow} + (V_w + V_o)P_w
% \end{equation}

% $A_{I_2}$ is the area of the interface without the particle. The change in the interface energy upon adsorption of the particle can be computed to obtain

% \begin{equation}
%     \Delta E_{IW} = A_o(\sigma_{po} - \sigma_{pw}) - A_p\sigma_{ow} + V_o(P_o - P_w)
% \end{equation}

% Substituting the Young-Laplace law for a spherical droplet, $P_w - P_o = \frac{2\sigma_{ow}}{R_d}$ where $R_d$ is the droplet size, for $(P_o - P_w)$ and the Young-Dupre contact angle
% expression, the following is obtained

% \begin{equation}
%     \Delta E_{IW} = \sigma_{ow}(A_o\cos{\theta_c} - A_p - \frac{2 V_o}{R_d})
% \end{equation}

% A similar construction can be done for the oil phase to obtain

% \begin{equation}
%     \Delta E_{IW} = -\sigma_{ow}(A_w\cos{\theta_c} + A_p - \frac{2 V_o}{R_d})
% \end{equation}

% After making appropriate geometrical substitutions for $V_o$ and $A_p$, the Piersanski model is recovered,

% \begin{equation}
%     \Delta E_{ads} = \pi r^2 \sigma_{ow} (1 - |\cos{\theta_c}|)^2
% \end{equation}

% From this expression, the adsorption energy can be seen to depend on the contact angle, surface tension and contact area of the particles. This indicates that larger, neutrally wetting
% particles adsorbing onto interfaces with high surface tension have a high driving force for adsorption. \cite{ngai_particle-stabilized_2015,reeves_particle-size_2015} This adsorption process 
% can be orders of magnitude larger than the thermal energy of the system, calculated as $k_b T$ where $T$ is the temperature of the system and $k_b$ is the Boltzmann constant. This results in 
% irreversible adsorption of particles at interfaces. \cite{binks_pickering_2001}

% In addition to the adsorption of particles on the interface, the contact angle of the particles determine the dispersed phase of an emulsion and its microstructure. 
% \cite{velankar_non-equilibrium_2015}
% Hydrophillic particles tend to prefer to form oil in water emulsions while hydrophobic particles prefer to form water in oil emulsions. \cite{ngai_particle-stabilized_2015}
% This occurs as the particles preferentially wet one phase over another, causing the interface to bend as the emulsion forms. Neutrally wetting particles are preferred in the formation of 
% bijels as no preferential curvature is imparted onto the bijels, fulfilling one of the requirements of a minimal surface in that the mean curvature must be zero. \cite{jinnai_interfacial_2001}

@@

As the library of available particle shapes continues to grow, it becomes increasingly important to understand how these anisotropic particles behave at fluid interfaces
\cite{wu_recent_2016, cavallaro_curvature-driven_2011}.
Unlike spherical particles, anisotropic particles introduce directional interactions and complex rotational dynamics which can significantly 
influence interfacial assembly and emulsion stability \cite{read_dimerization_2020, davies_dipolar_2015,morgan_understanding_2013}. 
Their shape can dictate not only their orientation at the interface but also how they interact, and jam, 
ultimately affecting the structure and mechanical properties of the resulting material \cite{hijnen_bijels_2015}.
This is relevant for systems like bijels, where the precise arrangement of particles at the interface affects how the bijel jams in place, generating new timescales or 
capillary bridging not characterized in bijels made with spheres \cite{gunther_timescales_2014,hijnen_bijels_2015,witt_bijel_2013}.

Anisotropic particles exhibit shape-dependent interfacial behavior due to their ability to induce complex multipolar capillary interactions. When adsorbed at a fluid-fluid interface, these 
particles distort the surrounding interface in order to maintain a mean curvature of zero, in accordance with the Young-Laplace equation. This distortion creates curvature gradients around the 
particle that interact with nearby particles or features on the interface, giving rise to anisotropic capillary forces \cite{loudet_capillary_2005, cheng_shape-anisotropic_2013}. 
Computational simulations further support this complexity, showing that ellipsoidal particles can adopt a variety of metastable orientations 
depending on initial conditions and perturbations encountered during adsorption or rearrangement at the interface \cite{gunther_lattice_2013, newton_influence_2014}.

@@

When looking at arrays of ellipsoidal particles at interfaces, curvature gradients control the dynamics of the system. Figure \ref{fig:curvature_gradient} illustrates the effect that
Cavallaro and co-workers identified. They showed how curvature gradients created at the edges of a 
square rod caused directed assembly of cylinders, with the driving force strong enough to overcome gravity \cite{cavallaro_curvature-driven_2011}. Other works have introduced how these features are 
particle shape dependent with greater control over the curvature gradient providing a means to control particle self assembly \cite{read_dimerization_2020, sharifi-mood_curvature_2015}.
Earlier work in the literature identified that ellipsoidal particles prefer to arrange side to side state while cylinders prefer to arrange tip to tip with the particle geometric smoothness playing 
a role in this preference \cite{botto_capillary_2012}. These differences in particle arrangement could open up new possibilities for tailoring interfacial assemblies and the subsequent differences 
in microstructure obtained in a bijel.

Ellipsoidal particles have been used as stabilizers in bijels, resulting in bijels that have smaller domains than those made with spherical particles at the same volume fraction. This was identified 
to arise from the greater surface area to volume ratio of ellipsoidal particles compared to spherical particles \cite{gunther_timescales_2014}. The evolution of the microstructure is also distinct 
to bijels stabilized with spheres, demonstrating two timescales at early and later timesteps \cite{gunther_timescales_2014}. The short timescale is during formation of the bijel, when particles 
are adsorbed onto the interface and rotate to align with the interface. At longer timescales, capillary interactions between particles lead to reordering of particles at the interface, leading 
to a different interfacial configuration and microstructure \cite{gunther_timescales_2014}.

@@

Rod like particles have also been used as stabilizers in experiments, identifying that the length scale of the pores of the bijels follows $ L \propto \frac{1}{\phi_p} $ even if the 
proportionality constant is different. Figure \ref{fig:rod_bijel_flippers} demonstrates bijels stabilized by rod like particles showing that particles flip out of the interface due to steric 
forces from adjacent particles. This flipping can provide structural support between the domains, providing the bijel with greater mechanical stability.
While the proportionality constant depends on particle geometry, this trend has been observed across various particle shapes 
\cite{hijnen_bijels_2015, madivala_exploiting_2009, gunther_timescales_2014, daware_emulsions_2015, loudet_capillary_2005, cheng_shape-anisotropic_2013}.

@@

Understanding the rheology of emulsions, particularly bijels, is essential because it directly impacts both their fabrication and practical use \cite{haase_situ_2016}.
Previous investigations have characterized that bijels are non-Newtonian fluids with a yield stress and shear thinning behavior \cite{macmillan_rheological_2019}.
A non-Newtonian fluid is one whose viscosity is not constant but varies with the applied shear rate or stress. In Newtonian fluids like water 
or air, there is a linear relationship between shear stress and shear rate, meaning viscosity remains the same regardless of how fast the fluid is sheared \cite{mezger_rheology_2020}. 
Non-Newtonian fluids, on the other hand, deviate from this behavior and may exhibit shear thinning, shear thickening, yield behavior or viscoelasticity with no fluid constrained to one behavior
\cite{mezger_rheology_2020}. Shear thinning and thickening here define that the viscosity increases at a slower and more rapid rate than the applied shear rate., while yield behavior implies that 
the viscosity is non-zero when no shear is applied \cite{mezger_rheology_2020}. Viscoelasticity appears when the shear response of the fluid is dependent upon the frequency of the applied shear, 
indicating that the fluid has both liquid and solid like characteristics \cite{mezger_rheology_2020}. These properties often appear in soft matter and multiphase systems where microstructure and 
particle interactions influence flow behavior.

@@

Suspensions and emulsions are examples of non-Newtonian fluids due to the presence of dispersed particles or droplets within a continuous phase
\cite{brader_nonlinear_2010, besseling_three-dimensional_2007, xu_relation_2013}.
In suspensions, the interaction  between solid particles—such as hydrodynamic forces, electrostatic repulsion, and steric hindrance—creates microstructural rearrangements under flow, leading to 
non-linear viscosity responses \cite{brader_nonlinear_2010, besseling_three-dimensional_2007}. Figure \ref{fig:suspension_shear} demonstrates the shear properties of an example suspension, 
demonstrating how shear thinning and shear thickening can occur in suspensions. At low shear, particles may aggregate or form a network that resists deformation, while higher shear rates can break 
down these structures, resulting in shear thinning. In dense systems, particles may experience contact-induced resistance or jamming, causing shear thickening \cite{brader_nonlinear_2010}.

Emulsions behave similarly with droplet deformation, coalescence, and interactions 
among droplets generating time-dependent and non-linear responses to shear. When particles are used to stabilize emulsions (e.g., in Pickering emulsions), their rigidity 
adds further resistance to flow and deformation, contributing to complex, often viscoelastic, rheological behavior.
Madivala et al. showed that Pickering emulsions stabilized with rod-like particles had enhanced stability under shear, with greater resistance observed as particle aspect ratio increased 
\cite{madivala_exploiting_2009}. They identified this improvement to be due to improved interfacial coverage of ellipsoidal particles at interfaces and the enhanced elasticity of the interface,
allowing for the droplet to elongate more before failure. These improvements also extend to disc like particles with other investigations into the resistance to shear of emulsions stabilized by
graphene nanosheets, showing that they were able to provide better resistance to shear than spherical particles \cite{imperiali_simple_2014}. This was due to the graphene nanosheets overlapping one
another during shear, creating greater resistance to the applied shear and a higher apparent viscosity. In this case too, the overlap of graphene nanosheets was shown to result in greater interfacial
elasticity inferred through surface tension measurements of droplets with graphene stabilizers \cite{sun_assembly_2013}.

@@

Several theoretical models have been developed to describe the behavior of non-Newtonian fluids under constant shear. The schematic representation of these models is shown in
Figure \ref{fig:shear_theory}. The power law Model is commonly used to represent shear thinning or 
thickening, with the equation \(\sigma = k \dot{\gamma}^n\), where \(\sigma\) is shear stress, \(\dot{\gamma}\) is shear rate, \(k\) is the consistency index, and \(n\) is the flow index 
\cite{mezger_rheology_2020}.
For \(n<1\), the fluid is shear thinning and for \(n>1\), shear thickening \cite{mezger_rheology_2020}.
The Bingham plastic model and Herschel-Bulkley model account for yield stress, with the latter being more general by 
including power-law behavior after yielding, defined as $\sigma = \sigma_y + k \dot{\gamma}^n$ \cite{mezger_rheology_2020}.
Shear models that include viscoelasticity have also been developed such as the  Maxwell and Kelvin-Voigt 
models \cite{mezger_rheology_2020}.
They are used when the fluid exhibits both solid- and liquid-like characteristics under stress, incorporating time-dependent strain responses and relaxation phenomena. These models help 
predict and quantify the behavior of complex fluids in real-world applications \cite{mezger_rheology_2020}.
